---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
<!-- badges: end -->

The *ffcr* package allows the construction of two families of transparent classification models: _fast-and-frugal trees_ and _tallying_ models. The book _Classification in the Wild: The Science and Art of Transparent Decision Making._[@katsikopoulos2021] describes these models, their applications and the algorithms to construct these in detail. 

A fast-and-frugal tree is a decision tree with a simple structure: one branch of each node exits tree, the other continues to the next node until the final node is reached. A tallying model gives pieces of evidence the same weight. The package implemts several methods for training these models, ranging from simple heuristics to computationyll more complex cross-entropy optimization (Rubinstein, 1999).


There exist another R package, [FFTrees](https://github.com/ndphillips/FFTrees) to train fast-and-frugal trees.


## Installation

You can install ffcr using from this GitHub page with

``` r
# install.packages("devtools")
devtools::install_github("marcusbuckmann/ffcr")
```
or download the Windows Binary _fftr_1.0.zip_ and install it from your hard drive.


## How to use the package


To illustrate the functionality of the package, we use the *Liver* data set [@ramana2011] that we obtained form the UCI machine learning repository [@dua2017]. It contains 579 patients of which 414 have a liver the condition and the other 165 do not. We predict which patient has a liver condition using medical measures and the age and gender of the people. 

We start with loading the package and the data.



```{r}
library(ffcr)
data(liver)
```



## Training fast-and-frugal trees

We start by training a fast-and-frugal tree on the Liver data set. If the first column in the data set is the class label, we can simply pass the data set as the first argument. 

```{r}
model <- fftree(liver, use_features_once = FALSE, method = "greedy", max_depth = 6)
```

The model object shows the structure of the trees and its performance on the data set.

```{r}
print(model)
```


To visualize the tree we use 
```{r, fig1, fig.height = 6, fig.width = 4, fig.align= "center"}
plot(model)
```

To make predictions according to a fast-and-frugal tree, we can use the `predict` function. 
```{r}
predict(model, newdata = liver[301:nrow(liver),], type = "metric")

```

## Training tallying models

To train tallying models, we use the *tally* function.



```{r}
model <- tally(liver, max_size = 6)
```

Please consult the vignette of the package for more details on the functionalities of the package and the book _Classification in the Wild_ for background information on fast-and-frugal trees and tallying models.

# References


Dua, Dheeru, and Casey Graff. 2017. “UCI Machine Learning Repository.” University of California, Irvine, School of Information; Computer Sciences. http://archive.ics.uci.edu/ml.

Katikopoulos, Konstantinos V., Özgür Şimşek, Marcus Buckmann, and Gerd Gigerenzer. 2020. Classification in the Wild: The Science and Art of Transparent Decision Making. MIT Press.

Ramana, Bendi Venkata, M Surendra Prasad Babu, and N. B. Venkateswarlu. 2011. “A Critical Study of Selected Classification Algorithms for Liver Disease Diagnosis.” International Journal of Database Management Systems 3 (2): 101–14.

Rubinstein, Reuven. 1999. “The Cross-Entropy Method for Combinatorial and Continuous Optimization.” Methodology and Computing in Applied Probability 1 (2): 127–90.

 


